---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: inline
---


## Loading and preprocessing the data

1. Load packages  
```{r echo=TRUE}
library(ggplot2)
library(dplyr)
library(Hmisc)
```

2. Download file and unzip  
```{r echo=TRUE}
wd <- getwd()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = paste0(wd,"/activity.zip"))
unzip("activity.zip")
```

3. load the 'activity.csv' file  
```{r echo=TRUE}
activity <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
1. total number of steps taken per day  
```{r echo=TRUE}
activity %>% group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps, na.rm = TRUE))
```
  
2. histogram of the total number of steps taken each day  
```{r echo=TRUE}
activity %>% group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps, na.rm = TRUE)) %>% 
  ggplot(aes(total_steps_per_day))+
  geom_histogram(binwidth = 1000)
```

  
3. mean and median of the total number of steps taken per day
```{r}
activity %>% group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps, na.rm = TRUE)) %>% 
  summarise(mean_total_steps_per_day = mean(total_steps_per_day))

activity %>% group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps, na.rm = TRUE)) %>% 
  summarise(median_total_steps_per_day = median(total_steps_per_day))
```
  


## What is the average daily activity pattern?  

1. time series plot  
```{r}
activity %>% group_by(interval) %>% 
  summarise(avg_step = mean(steps, na.rm = TRUE)) %>% 
  ggplot(aes(x = interval, y = avg_step))+
  geom_line()+
  labs(x = "5-minute interval", y = "average number of steps", title = "Average daily activity pattern")
```

2. 5-minute interval contains the maximum number of steps
```{r}
activity %>% group_by(interval) %>% 
  summarise(avg_step = mean(steps, na.rm = TRUE)) %>% 
  arrange(desc(avg_step)) %>% 
  head(1)
```



## Imputing missing values

1. total number of missing values
```{r}
summary(activity)
activity %>% filter(is.na(steps)) %>% count()
```

2. filling in all of the missing values with mean
```{r}
activity$steps_impute <- impute(activity$steps, fun=mean)

```

3. histogram of the total number of steps taken each day
```{r}
activity %>% group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps_impute, na.rm = TRUE)) %>% 
  ggplot(aes(total_steps_per_day))+
  geom_histogram(binwidth = 1000)
```

mean and median  
```{r}
activity %>% group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps_impute, na.rm = TRUE)) %>% 
  summarise(mean_total_steps_per_day = mean(total_steps_per_day))

activity %>% group_by(date) %>% 
  summarise(total_steps_per_day = sum(steps_impute, na.rm = TRUE)) %>% 
  summarise(median_total_steps_per_day = median(total_steps_per_day))
```



## Are there differences in activity patterns between weekdays and weekends?

1. new factor variable – “weekday” and “weekend”  
```{r}
Sys.setlocale("LC_TIME", "English")
activity_week <- activity %>% mutate(days = weekdays(as.Date(date)),
                    weekday = ifelse(days %in% c("Sunday", "Saturday"), "weekend", "weekday"))
```

2. time series plot – “weekday” and “weekend” 
```{r}
activity_week %>% group_by(weekday,interval) %>% 
  summarise(avg_step = mean(steps, na.rm = TRUE)) %>% 
  ggplot(aes(x = interval, y = avg_step, color = weekday))+
  geom_line()+
  labs(x = "5-minute interval", y = "average number of steps", 
       title = "Average daily activity pattern : “weekday” vs  “weekend”")+
  facet_grid(weekday ~.)
```

