---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

1. Load packages  
```{r}
library(dplyr)
library(ggplot2)
```

2. Download file and unzip  
```{r}
wd <- getwd()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = paste0(wd,"/activity.zip"))
unzip("activity.zip")
```

3. load the 'activity.csv' file  
```{r}
activity <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
1. total number of steps taken per day  
```{r}
activity %>% group_by(date) %>% 
  summarize(total_steps_per_day = sum(steps, na.rm = TRUE))
```
  
2. histogram of the total number of steps taken each day  
```{r}
activity %>% group_by(date) %>% 
  summarize(total_steps_per_day = sum(steps, na.rm = TRUE)) %>% 
  ggplot(aes(total_steps_per_day))+
  geom_histogram(binwidth = 1000)
```
  
3. mean and median of the total number of steps taken per day
```{r}
activity %>% group_by(date) %>% 
  summarize(total_steps_per_day = sum(steps, na.rm = TRUE)) %>% 
  summarize(mean_total_steps_per_day = mean(total_steps_per_day))

activity %>% group_by(date) %>% 
  summarize(total_steps_per_day = sum(steps, na.rm = TRUE)) %>% 
  summarize(median_total_steps_per_day = median(total_steps_per_day))
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
